Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-22 paper.tex
Minor changes. Added hlines to Table 2

\begin{document}
\maketitle
%
\begin{abstract}
Given how common IoT devices that use constrained resources are becoming today, the need of the hour is communication protocols which can operate securely under such limitations.
%
For a few years, the Internet Engineering Task Force (IETF) has been working to standardize \mEdhoc{}, an authenticated key establishment protocol for such constrained IoT devices.
%
The first version of \mEdhoc{} was proposed in 2016.
%
In 2018, Bruni et al.~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of \mEdhoc{}, which had only two key establishment methods.  
%
%By now, the protocol has been fleshed out much more, and has been augmented with multiple new key establishment methods.
By 2021, the protocol had been fleshed out much more, with multiple new key establishment methods, and this version was formally analyzed using the \mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
%
Here, we build on this work, and use \mTamarin{} to formally analyze the key establishment methods in the current version of \mEdhoc{}, as well as discuss some ramifications of the choices made while designing the protocol.
\vnote{We need to see what more we can say here, about the new stuff we're potentially adding}
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}
\input{intro}

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Error handling}}
\label{sec:errorHandling}
Section 6 of the current \mEdhoc{} specification~\cite{draft-ietf-lake-edhoc-15} states that error messages can be sent at any time and by any party.
%
There are three types of error messages: type 0 is used to represent success, type 1 represents a generic error message used for debugging purposes, and type 2 represents the failure to negotiate a common ciphersuite. 
%
Type 2 messages carry a list of supported ciphersuites by the responder.
%
The contents and semantics of type 1 messages are dependent on the implementation.
%
Consequently, type 1 error messages provide a generic message passing mechanism, albeit without a predefined semantics.
%
We now discuss some issues with this specification of error messages.

@@

\subsection{Unspecified semantics for debugging messages}
Section 6.2 gives an example of sending a type 1 error message with the information string "Method not supported".
%
Note that \mEdhoc{} does not include a mechanism to support the negotiating of the connection method.
%
The initiator selects the method as part of the first message, and if the responder replies with a type 1 error message along with the information string "Method not supported", the standard does not tell an implementation how to act upon this information string for negotiation purposes.
%
An adversary can fake an error message, and there is no algorithm that is specified to ensure that a common available authentication mechanism is always selected.
%
In particular, note that Section 5.1 in the specification states that if processing fails for some reason, then "typically" an error message is sent, processing is terminated and protocol state is discarded. 
%
But negotiation needs us to remember state.

With respect to ciphersuite negotiation, Section 5.2.2 of the specification also states ``The Initiator MUST NOT change the supported cipher suites and
the order of preference in SUITES\_I based on previous error messages.'' 
%
However, \mSuites is sent in the clear, and therefore can be modified by the adversary.
%
It is important to consider what happens if the order of preference in \mSuites is changed either by the adversary, or by the initiator themselves for some other reason.
%
It might be helpful to include a flag in \mMsgone to indicate to the responder
that a (modified) ciphersuite ordering depends on that from a previous
execution.

%-------------------------------------------------------------------------- sec
\section{\uppercase{Improvements to the specification}}
\label{sec:discussion}
The \mEdhoc{} \mSpec{} has many points where it can be improved.
%
We discovered a few of these issues in~\cite{Norr21}, but in this work,
we also discuss a couple of new issues.
%
We communicated the points discussed in~\cite{Norr21} to the authors of the
\mSpec{}.
%
Here we also list the protocol developers' response to being told of these
ambiguities in the \mSpec{}.

@@

A straight forward check that the received \mIdcredr{} matches the expected
identity thwarts the problem.
%
We included this check in our model.
%
We communicated this to the authors of \mEdhoc{}, who in the latest version of
the \mSpec{}, added that \mEdhoc{} makes \mIdcredr{} available to the
application, and added a new appendix D describing credential validations that
the application must take care of.
%

@@

Section 8.7 of the latest version of the \mEdhoc{}
\mSpec{}~\footnote{draft-ietf-lake-edhoc-15}
gives two recommendations for countering DoS attacks.
%
First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS
countermeasures is given: checking the return address upon receipt of the first
message.
%
Second, a recommendation is given that applications may try to determine (in
some way) that a seemingly valid message is in fact probably a forgery and
should be ignored or cause the protocol to abort.
%

@@

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
In addition to unclear contexts in which
\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail about
adversary capabilities.
%
The four different cryptographic cores included in \mEdhoc{} are based on
academic protocols which are designed to work with highly specific but
potentially different adversary models.
%
%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, which
%explicitly separates long-term key storage from that of the ephemeral keys and
%session state.
%%
%This fits in well with our use of TEEs.
Since \mSigma{} cannot protect against compromised ephemeral keys, the authors
of \mEdhoc{} felt that considering the compromise of ephemeral keys separately
from that of long-term keys was not required~\cite{personalCommunication}.
%
The reason was presumably based on the fact that the \mSigSig{} method was
modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.~\cite{Norr21}
%
However, this only holds true for situations where one is only interested in the
session key confidentiality for an ongoing session.
%
Secure modules, in addition to being able to store long-term keys, are useful in
other ways too.
%
They also provide weak Post-Compromise Security (PCS) guarantees as discussed
above.
%
Upon discussion with the authors, they included recommendations on how to store
long-term keys, and how to perform operations on these keys inside a secure
module.
%

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusions}
\label{sec:newdrafts}
We formally modeled all four
methods of the \mEdhoc{} \mSpec{} using \mTamarin.
%
We formulated several important security properties and identified precise
adversary models in which we verified these.
%
The properties are shown in Table~\ref{tab:props}.
%
Mutual injective agreement covers the parameters $S_P$:
responder identity, roles, session key material (except for $P_I$ when
initiator uses the \mStat{} authentication
method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
%
The responder in addition is ensured agreement on the initiators identity and
$P_I$, i.e., on the set $S_F$.
%
Implicit agreement covers all previously mentioned parameters for both peers.
%
All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
256 GiB of memory for the scheduled job (the machine has a total of
384 GiB installed).
%
Time measurements are rounded to the nearest second and we report both
real and user times (i.e. time to complete the task and total CPU time
spent for the task).
%
Mutual entity authentication, UKS- and KCI resistance can be inferred
from the verified properties.
\knote{In Section 3.3 we say that injective agreement does \emph{not} hold for
    all methods. It is a little contradictory to the table below where we
    exclude $P_I$. It would be better to show it as not holding when $P_i$ is
    included (both for consistency and for clarity).
}
%
\begin{table*}[t]
        \centering
          \label{tab:props}
          \resizebox{\textwidth}{!}{%
            \begin{tabular}{l|c|c|c|c}
              \textbf{Lemma $\backslash$ Method} & \textbf{SIG-SIG} & \textbf{SIG-STAT} & \textbf{STAT-SIG} & \textbf{STAT-STAT}\\
              \hline
              \textbf{Injective agreement / I} & $S_F$ & $S_F$ & $S_P\ddagger$ & $S_P\ddagger$\\
              \textbf{Injective agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 1m21s / 35m36s & 18m16s / 499m2s\\
              \textbf{Implicit agreement / I} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Implicit agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s $\dagger$ & 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
              \textbf{Session key secrecy} & \cm & \cm & \cm & \cm\\
              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 156m26s & 9m47s / 232m2s\\
              \textbf{Secrecy of AD3} & \cm & \cm & \cm & \cm\\
              \textbf{Integrity of AD3} & \cm & \cm & \cm & \cm\\
              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 3m35s / 85m12s & 45m6s / 1029m1s\\
            \end{tabular}}\\[1em]

          \caption{Verified properties. $S_P$ contains roles,
            responder identity, session key material (excluding
            $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
            the initiator identity, and $P_I$.\\
            $\dagger$: For SIG-STAT and STAT-STAT implicit
            authentication using the full XOR model the verification
            was killed at 14246m43s and 2057m47s of computation time,
            respectively. The time reported is for verification under
            the simplified XOR modeling of message 2.\\
            $\ddagger$: When the initiator is using STAT mode we
            cannot get explicit agreement on the intiator's own key
            material, therefore in these two cases we check injective
            agreement for all key material except $P_I$.\\
          }
        % \begin{tabular}{|l|c|c|c|c|}
        %         \hline
        %         & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
        %         \hline
        %         Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
        %         Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         PFS for session key material & \cm & \cm & \cm & \cm\\
        %         \hline
        % \end{tabular}
\end{table*}

@@

%\subsection{Newer Versions of the Specification}
%\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} version as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
There are newer versions and our analysis applies to at least the version from
February 2021~\cite{latest-ietf-lake-edhoc-05}.
%
The latest version is \emph{draft-ietf-lake-edhoc-15}
%
\knote{we are already citing too many different versions, so I don't think we
    should add this to the bib. Potentially, we should tomorrow instead remove
    and just mention some of the later versions.}
%

>>>>

\begin{document}
\maketitle
%
\begin{abstract}
Given how common IoT devices that use constrained resources are becoming 
today, the need of the hour is communication protocols which can operate 
securely under such limitations.
%
For a few years, the Internet Engineering Task Force (IETF) has been working 
to standardize \mEdhoc{}, an authenticated key establishment protocol for 
such constrained IoT devices.
%
The first version of \mEdhoc{} was proposed in 2016.
%
In 2018, Bruni et al.~\cite{DBLP:conf/secsr/BruniJPS18} used the \mProverif{} 
tool~\cite{DBLP:conf/csfw/Blanchet01} to formally analyze an early version of 
\mEdhoc{}, which had only two key establishment methods.  
%
%By now, the protocol has been fleshed out much more, and has been 
%augmented with multiple new key establishment methods.
By 2021, the protocol had been fleshed out much more, with multiple new key 
establishment methods, and this version was formally analyzed using the 
\mTamarin{} prover~\cite{DBLP:conf/cav/MeierSCB13} in~\cite{Norr21}.
%
Here, we build on this work, by modifying the model, adding new properties 
that we analyze, and discussing some of the aspects of the latest \mEdhoc{} 
specification. In particular, we extend the modelling with trusted execution 
environments (TEEs), modify the way we model \mXor{} encryption, and 
verify weak post-compromise security (PCS) as well as the secrecy and 
integrity of some additional data used as part of the protocol.
\end{abstract}
%
%-------------------------------------------------------------------------- sec
\section{\uppercase{Introduction}}
\label{sec:introduction}
\input{intro}

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Error handling}}
\label{sec:errorHandling}
Section 6 of the current \mEdhoc{} 
specification~\cite{draft-ietf-lake-edhoc-15} 
states that error messages can be sent at any time and by any party.
%
There are three types of error messages: type 0 is used to represent success, 
type 1 represents a generic error message used for debugging purposes, and 
type 2 represents the failure to negotiate a common ciphersuite. 
%
Type 2 messages carry a list of supported ciphersuites by the responder.
%
The contents and semantics of type 1 messages are dependent on the 
implementation.
%
Consequently, type 1 error messages provide a generic message passing 
mechanism, albeit without a predefined semantics.
%
We now discuss some issues with this specification of error messages.
%
In this section, we use ``specification'' to mean the latest specification, as 
in~\cite{draft-ietf-lake-edhoc-15}.

@@

\subsection{Unspecified semantics for debugging messages}
Section 6.2 gives an example of sending a type 1 error message with the 
information string ``Method not supported".
%
Note that \mEdhoc{} does not include a mechanism to support the 
negotiating of the connection method.
%
The initiator selects the method as part of the first message, and if the 
responder replies with a type 1 error message along with the information 
string ``Method not supported", the standard does not tell an 
implementation how to act upon this information string for negotiation 
purposes.
%
An adversary can fake an error message, and there is no algorithm that is 
specified to ensure that a common available authentication mechanism is 
always selected.
%
In particular, note that Section 5.1 in the specification states that if processing 
fails for some reason, then ``typically" an error message is sent, processing is 
terminated and protocol state is discarded. 
%
But negotiation needs us to remember state.

With respect to ciphersuite negotiation, Section 5.2.2 of the specification also 
states ``The Initiator MUST NOT change the supported cipher suites and
the order of preference in SUITES\_I based on previous error messages.'' 
%
However, \mSuites is sent in the clear, and therefore can be modified by the 
adversary.
%
It is important to consider what happens if the order of preference in 
\mSuites{} is changed either by the adversary, or by the initiator themselves 
for some other reason.
%
It might be helpful to include a flag in \mMsgone{} to indicate to the 
responder
that a (modified) ciphersuite ordering depends on that from a previous
execution.

%-------------------------------------------------------------------------- sec
\section{\uppercase{Improvements to the specification}}
\label{sec:discussion}
The \mEdhoc{} \mSpec{} has many points where it can be improved.
%
We discovered a few of these issues in~\cite{Norr21}, but in this work,
we also discuss a couple of new issues.
%
We communicated the points discussed in~\cite{Norr21} to the authors of the
\mSpec{}.
%
Here, we also list the protocol developers' response (if any) to being told of these
ambiguities in the \mSpec{}.

@@

A straight forward check that the received \mIdcredr{} matches the expected
identity thwarts the problem.
%
We included this check in our model.
%
We communicated this to the authors of \mEdhoc{}, who in the latest version 
of
the \mSpec{}, added that \mEdhoc{} makes \mIdcredr{} available to the
application, and added an appendix D listing credential validations 
that
the application must take care of.
%

@@

Section 8.7 of the latest version of the \mEdhoc{}
\mSpec{}~\cite{draft-ietf-lake-edhoc-15}
gives two recommendations for countering DoS attacks.
%
First, \mEdhoc{} relies on lower layers to mitigate DoS, and an example of DoS
countermeasures is given: checking the return address upon receipt of the 
first
message.
%
Second, a recommendation is given that applications may try to determine (in
some way) that a seemingly valid message is in fact probably a forgery and
should be ignored or cause the protocol to abort.
%

@@

%------------------------------------------------------------------------- sub
\subsection{Unclear Security Model}
In addition to unclear contexts in which
\mEdhoc{} can be used, the \mSpec{} also does not provide enough detail 
about
adversary capabilities.
%
The four different cryptographic cores included in \mEdhoc{} are based on
academic protocols which are designed to work with highly specific but
potentially different adversary models.
%
%For example, \mOptls{}, which essentially has the same core as \mStat{}, is
%shown to be secure in the CK model~\cite{DBLP:conf/crypto/CanettiK02}, 
%which
%explicitly separates long-term key storage from that of the ephemeral keys 
%and
%session state.
%%
%This fits in well with our use of TEEs.
Since \mSigma{} cannot protect against compromised ephemeral keys, the 
authors
of \mEdhoc{} felt that considering the compromise of ephemeral keys 
separately
from that of long-term keys was not required~\cite{personalCommunication}.
%
The reason was presumably based on the fact that the \mSigSig{} method was
modeled closely on the \mSigmaI{} variant of \mSigma{}, and that it would be
preferable to obtain a homogeneous security level among the \mEdhoc{}
methods.~\cite{Norr21}
%
However, this only holds true for situations where one is only interested in 
the
session key confidentiality for an ongoing session.
%
Secure modules, in addition to being able to store long-term keys, are useful 
in
other ways too.
%
They also provide weak Post-Compromise Security (PCS) guarantees as 
discussed
above.
%
Upon discussion with the authors, they included recommendations on how to 
store
long-term keys, and how to perform operations on these keys inside a secure
module.
%

@@

%-------------------------------------------------------------------------- sec
\section{\uppercase{Conclusions and Future Work}}
\label{sec:conclusion}
%\label{sec:newdrafts}
We formally modeled all four
methods of the \mEdhoc{} \mSpec{} using \mTamarin.
%
We formulated several important security properties and identified precise
adversary models in which we verified these.
%
The properties are shown in Table~\ref{tab:props}.
%
Mutual injective agreement covers the parameters $S_P$:
responder identity, roles, session key material (except for $P_I$ when
initiator uses the \mStat{} authentication
method), context identifiers \mCi{} and \mCr, and cipher suites \mSuites.
%
The responder in addition is ensured agreement on the initiators identity and
$P_I$, i.e., on the set $S_F$.
%
Implicit agreement covers all previously mentioned parameters for both 
peers.
%
All measurements were taken on a server with 2 Intel(R) Xeon(R) Gold
6242 CPU @ 2.80GHz with a total of 32 cores / 64 threads and reserved
256 GiB of memory for the scheduled job (the machine has a total of
384 GiB installed).
%
Time measurements are rounded to the nearest second and we report both
the real time (time to complete the task) and user time (total CPU time
spent for the task).
%
Mutual entity authentication, UKS- and KCI-resistance can be inferred
from the verified properties.
%

\begin{table*}[t]
        \centering
          \resizebox{\textwidth}{!}{%
            \begin{tabular}{|l|c|c|c|c|}
            \hline
              \textbf{Lemma $\backslash$ Method} & \textbf{SIG-SIG} & 
\textbf{SIG-STAT} & \textbf{STAT-SIG} & \textbf{STAT-STAT}\\
              \hline
              \textbf{Injective agreement / I} & $S_F$ & $S_F$ & $S_P\ddagger$ & 
$S_P\ddagger$\\
              \textbf{Injective agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Time (real/user)} & 0m27s / 11m17s & 9m27s / 239m49s & 
1m21s / 35m36s & 18m16s / 499m2s\\
              \textbf{Implicit agreement / I} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Implicit agreement / R} & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
              \textbf{Time (real/user)} & 2m37s / 65m40s & 5m28s / 131m12s 
$\dagger$ 
& 8m54s / 212m21s & 7m37s / 188m46s $\dagger$\\
              \textbf{Session key secrecy} & \cm & \cm & \cm & \cm\\
              \textbf{Time (real/user)} & 0m41s / 16m4s & 1m52s / 45m27s & 6m31s / 
156m26s & 9m47s / 232m2s\\
              \textbf{Secrecy of AD3} & \cm & \cm & \cm & \cm\\
              \textbf{Integrity of AD3} & \cm & \cm & \cm & \cm\\
              \textbf{Time (real/user)} & 1m40s / 38m57s & 5m53s / 138m43s & 
3m35s / 85m12s & 45m6s / 1029m1s\\
			\hline
            \end{tabular}}\\[1em]
          \caption{Verified properties. $S_P$ contains roles,
            responder identity, session key material (excluding
            $P_I$), \mCi, \mCr, and \mSuites. $S_F$ is $S_{P}$,
            the initiator identity, and $P_I$.\\
            $\dagger$: For SIG-STAT and STAT-STAT implicit
            authentication using the full XOR model the verification
            had to be abandoned at 14246m43s and 2057m47s of computation time,
            respectively. The time reported is for verification under
            the simplified XOR modeling of message 2.\\
            $\ddagger$: When the initiator is using STAT mode we
            cannot get explicit agreement on the intiator's own key
            material, therefore in these two cases we check injective
            agreement for all key material except $P_I$.\\
          }
                    \label{tab:props}
        % \begin{tabular}{|l|c|c|c|c|}
        %         \hline
        %         & \mSigSig & \mSigStat & \mStatSig & \mStatStat \\
        %         \hline
        %         Injective agreement for I & $S_F$ & $S_F$ & $S_P$ & $S_P$\\
        %         Injective agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         Implicit agreement for I & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         Implicit agreement for R & $S_F$ & $S_F$ & $S_F$ & $S_F$\\
        %         PFS for session key material & \cm & \cm & \cm & \cm\\
        %         \hline
        % \end{tabular}
\end{table*}

@@

\subsection{A Note About the Version of \mEdhoc}
\label{sec:newdrafts}
In this work, we have analyzed the \mEdhoc{} specification as of July
2020~\cite{our-analysis-selander-lake-edhoc-00}.
%
While our formal analysis applies to at least the version from February 2021, there are newer versions.
%
The latest version is from July 2022~\cite{draft-ietf-lake-edhoc-15}, which, among other things, changes the key derivation function.
%
However, we do refer to various aspects (error handling, denial of service etc) of this latest version, in Sections~\ref{sec:errorHandling} and \ref{sec:discussion}.

%

\knote{we 
%are already citing too many different versions, so I don't think we
    should add this to the bib. Potentially, we should tomorrow instead remove
    and just mention some of the later versions.}
\vnote{We do need to cite the latest one, just so people can look it up. It would 
be more sensible to remove the citation for the one from February 2021, and 
instead be more concrete about which version of EDHOC is the latest one that 
our formal analysis applies to.}
%

### 2022-07-22 protocol.tex
Minor changes. Added hlines to Table 2

At the end of a successful run of the protocol, the session key material is
established as \mSessKey{}, which we define as a set of various keys.
%
This set always includes \mGxy{}, and if the initiator (resp. the responder) uses the \mStat{} authentication method, then it also includes $P_{I}$ (resp. $P_{R}$).
%
We discuss which material should be included in \mSessKey{} in more detail in
the formal model in Section~\ref{sec:formalization} and the consequences of
various choices in Section~\ref{sec:sessionKeyMaterial}.
%
Once \mSessKey{} has been established, an \mHkdf{}-based key exporter named
\mEdhoc-Exporter extracts the keys required by the security protocol.
%

@@

In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution under
the \mSigStat{} method.
%
The figure describes in detail the various message patterns, operations and
key derivations used to construct these messages.
\vnote{Do we still have the other method diagrams that we wrote out for the original paper but did not include?}
\knote{I don't think we have space for that.}

\begin{figure}[ht]
\centering
\scalebox{.7}{
\tikzset{>=latex, every msg/.style={draw=thick}, every node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.5) -- (0,-14.8);
    \draw [very thick] (9, -0.5) -- (9,-14.8);
    \node[below=0.5em of ini,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ \mADone,\ \mADthree
    \end{array}
    $};
    \node (res) at (9,0) {Responder};
    \node[below=0.5em of res,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{R},\ \mPub{R}\rangle, \ \mIdcredr,\ \mADtwo
    \end{array}$};
    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ \mGx\rangle$};
    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, \mADone};
    \action{8em}{res}{$
      \begin{array}{c}
          \text{Generates } \mCr,\ \langle\mY{},\ \mGy\rangle\\
          \ \ P_e = \mDH(\mY,\ \mGx{})\\
          \ \ P_R = \mDH(\mPriv{R},\ \mGx{})\\
        \mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle)\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, \mADtwo \rangle; \textrm{``\phantom{}''}) \\
        \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
      \end{array}$};
    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ \mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
    \action{25em}{ini}{$
      \begin{array}{c}
        %\mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle) \
        \ P_e = \mDH(\mX,\ \mGy{})\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        %\mKtwoe = \mHkdfExpand(\mPRKtwo,\mTHtwo)\\
        %\mGrx = \mCredr^{x} \\
        \ \ P_R = \mDH(\mX,\ \mPub{R})\\
        \mPRKfour = \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        %\mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, \mADthree \rangle, \mMacthree \rangle)
      \end{array}$};
    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; \mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
    \action{40em}{res}{$
    \begin{array}{c}
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKthree, \mTHthree) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHthree)
    \end{array}$};
    \draw [line width=1mm] (-2,-14.8) -- (2,-14.8);
    \draw [line width=1mm] (7,-14.8) -- (11,-14.8);
    \end{tikzpicture}
}
    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$ denotes a tuple, and the hash function \mHash{} is as established in the ciphersuite \mSuites{}.~\cite{Norr21}}
\label{fig:edhocsigstat}
\end{figure}

>>>>

At the end of a successful run of the protocol, the session key material is
established as \mSessKey{}, which we define as a set of various keys.
%
This set always includes \mGxy{}, and if the initiator (resp. the responder) uses 
the \mStat{} authentication method, then it also includes $P_{I}$ (resp. 
$P_{R}$).
%
We 
%discuss which material should be included in \mSessKey{} in more detail 
in
the formal model in Section~\ref{sec:formalization} and the consequences of
various choices in Section~\ref{sec:sessionKeyMaterial}.
%
Once \mSessKey{} has been established, an \mHkdf{}-based key exporter 
named
\mEdhoc-Exporter extracts the keys required by the security protocol.
%

@@

In Figure~\ref{fig:edhocsigstat} we show an example of protocol execution 
under
the \mSigStat{} method.
%
The figure describes in detail the various message patterns, operations and
key derivations used to construct these messages.
%\vnote{Do we still have the other method diagrams that we wrote out for the 
%original paper but did not include?}
%\knote{I don't think we have space for that.}
%
\begin{figure}[ht]
\centering
\scalebox{.7}{
\tikzset{>=latex, every msg/.style={draw=thick}, every 
node/.style={fill=none,text=black}}
\begin{tikzpicture}
    \node (ini) at (0, 0) {Initiator};
    \draw [very thick] (0, -0.5) -- (0,-14.8);
    \draw [very thick] (9, -0.5) -- (9,-14.8);
    \node[below=0.5em of ini,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{I},\ \mPub{I}\rangle,\ \mIdcredi,\ \mIdcredr,\ 
\mADone,\ \mADthree
    \end{array}
    $};
    \node (res) at (9,0) {Responder};
    \node[below=0.5em of res,fill=white] {$
    \begin{array}{c}
        \text{Knows}\ \langle\mPriv{R},\ \mPub{R}\rangle, \ \mIdcredr,\ \mADtwo
    \end{array}$};
    \action{3.7em}{ini}{Generates $\mMethod,\ \mSuites,\ \mCi,\ \langle\mX{},\ 
\mGx\rangle$};
    \msg{7em}{ini}{res}{\mMsgone: \mMethod, \mSuites, \mGx, \mCi, 
\mADone};
    \action{8em}{res}{$
      \begin{array}{c}
          \text{Generates } \mCr,\ \langle\mY{},\ \mGy\rangle\\
          \ \ P_e = \mDH(\mY,\ \mGx{})\\
          \ \ P_R = \mDH(\mPriv{R},\ \mGx{})\\
        \mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle)\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        \mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mMactwo = \mAead(\mKtwom; \langle \mIdcredr, \mTHtwo, \mCredr, 
\mADtwo \rangle; \textrm{``\phantom{}''}) \\
        \mKtwoe = \mHkdfExpand(\mPRKtwo, \mTHtwo)
      \end{array}$};
    \msg{24em}{res}{ini}{\mMsgtwo: \mCi, \mGy, \mCr, $\overbrace{\mKtwoe\ 
\mXor\ \langle \mIdcredr, \mMactwo, \mADtwo \rangle}^{\mCipher}$};
    \action{25em}{ini}{$
      \begin{array}{c}
        %\mTHtwo = \mHash(\mMsgone, \langle \mCi, \mGy, \mCr \rangle) \
        \ P_e = \mDH(\mX,\ \mGy{})\\
        \mPRKtwo = \mHkdfExtract(\textrm{``\phantom{}''}, P_e) \\
        %\mKtwoe = \mHkdfExpand(\mPRKtwo,\mTHtwo)\\
        %\mGrx = \mCredr^{x} \\
        \ \ P_R = \mDH(\mX,\ \mPub{R})\\
        \mPRKfour = \mPRKthree = \mHkdfExtract(\mPRKtwo, P_R) \\
        %\mKtwom = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHtwo) \\
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKfour, \mTHthree) \\
        \mMacthree = \mAead(\mKthreem; \langle \mIdcredi, \mTHthree, 
\mCredi, \mADthree \rangle; \textrm{``\phantom{}''}) \\
        \mSigthree = \mSign{I}(\langle \langle \mIdcredi, \mTHthree, \mCredi, 
\mADthree \rangle, \mMacthree \rangle)
      \end{array}$};
    \msg{38.5em}{ini}{res}{$\mMsgthree: \mCr, \mAead(\mKthreeae; 
\mTHthree; \langle \mIdcredi, \mSigthree, \mADthree \rangle$)};
    \action{40em}{res}{$
    \begin{array}{c}
        \mTHthree = \mHash(\mTHtwo, \mCipher, \mCr)\\
        \mKthreem = \mHkdfExpand(\mPRKthree, \mTHthree) \\
        \mKthreeae = \mHkdfExpand(\mPRKthree, \mTHthree)
    \end{array}$};
    \draw [line width=1mm] (-2,-14.8) -- (2,-14.8);
    \draw [line width=1mm] (7,-14.8) -- (11,-14.8);
    \end{tikzpicture}
}
    \caption{The \mSigStat{} method for \mEdhoc{}. $\langle\cdot\rangle$ 
denotes a tuple, and the hash function \mHash{} is as established in the 
ciphersuite \mSuites{}.~\cite{Norr21}}
\label{fig:edhocsigstat}
\end{figure}
